\bf Tearing Mode Dynamics in Tokamak Plasmas

### 2021-11-16 Documentation/TearingModeTheory.tex
Latest version

If we assume that $\delta_e\sim \zeta_e\sim \mu\ll 1$ then the dominant terms in the electron energy conservation equation
(\ref{e3.151c}) yield
\begin{equation}
\nabla\cdot{\bf q}_{T_e}\simeq \nabla\cdot(-\kappa_{\parallel\,e}\,\nabla_\parallel T_e)\simeq 0,
\end{equation}
which implies that 
\begin{equation}
{\bf B}\cdot\nabla T_e=0. \label{e75}
\end{equation}
In other words, the parallel electron heat conductivity in a tokamak plasma is usually
sufficiently large that it forces the electron temperature to be constant on magnetic flux-surfaces. 
Likewise, if we assume that $\delta_i\sim \zeta_i\sim \mu\ll 1$ then the dominant terms in the ion energy conservation equation
(\ref{e3.152c}) yield
\begin{equation}
\nabla\cdot{\bf q}_{T_i}\simeq \nabla\cdot(-\kappa_{\parallel\,i}\,\nabla_\parallel T_i)\simeq 0,
\end{equation}
which implies that 
\begin{equation}
{\bf B}\cdot\nabla T_i=0.\label{e77}
\end{equation}
In other words, the parallel ion heat conductivity in a tokamak plasma is usually
sufficiently large that it forces the ion temperature to be constant on magnetic flux-surfaces.

 The dominant terms in the electron and ion momentum conservation equations,
(\ref{e3.151b}) and (\ref{e3.152b}), respectively, yield
\begin{align}\label{e78}
\nabla p_e + n\,e\,({\bf E}+{\bf V}_e\times {\bf B}) &\simeq 0,\\[0.5ex]
\nabla p_i - n\,e\,({\bf E}+{\bf V}_i\times {\bf B}) &\simeq 0.\label{e79}
\end{align}
The sum of the preceding two equations gives
\begin{equation}\label{e80}
 {\bf j}\times {\bf B} \simeq\nabla p,
\end{equation}
where $p=p_e+p_i$. In other words, a conventional tokamak plasma exists in a state in which the magnetic force density exactly balances the  total scalar pressure force density. 
It follows from the previous equation that 
\begin{equation}
{\bf B}\cdot\nabla p\simeq 0.
\end{equation}
In other words, the total pressure in a conventional tokamak plasma is constant on magnetic flux-surfaces. 
 However, given that $p=n\,(T_e+T_i)$, and making use of Eqs.~(\ref{e75}) and (\ref{e77}), we deduce that 
\begin{equation}
{\bf B}\cdot\nabla n\simeq 0.
\end{equation}
In other words, the electron number density in a conventional tokamak plasma is also  constant on
magnetic flux-surfaces.

@@

\subsection{Solution in Presence of Perfectly-Conducting Wall}\label{perfect}
Suppose that the plasma occupies the region $0\leq r\leq a$, where $a$ is the plasma {\em minor radius}. It follows that
$J_z(r)=0$ for $r>a$. Let the plasma be surrounded by a concentric, rigid, radially-thin, perfectly-conducting, wall of radius $r_w>a$. 
(By definition, $a\ll R_0$ in a large aspect-ratio tokamak. In most circumstances, the wall represents the metal vacuum vessel that surrounds  the plasma.) An appropriate
physical solution of the cylindrical tearing mode equation takes the separable form
\begin{equation}\label{e99}
\psi(r,t)= {\mit\Psi}_s(t)\,\hat{\psi}_s(r),
\end{equation}
where the real function $\hat{\psi}_s(r)$ is a solution of 
\begin{equation}\label{e100}
\frac{d^2\hat{\psi}_s}{dr^2} + \frac{1}{r}\,\frac{d\hat{\psi}_s}{dr}-\frac{m^2}{r^2}\,\hat{\psi}_s - \frac{J_z'\,\hat{\psi}_s}{r\,(1/q-n/m)}= 0
\end{equation}
that satisfies
\begin{align}\label{e101}
\hat{\psi}_s(0) &= 0,\\[0.5ex]
\hat{\psi}_s(r_s) &= 1,\label{e102}\\[0.5ex]
\hat{\psi}_s(r\geq r_w) &= 0.\label{e103}
\end{align}

@@

The complex quantity
\begin{equation}\label{e109}
{\mit\Delta\Psi}_s = \left[r\,\frac{\partial \psi}{\partial r}\right]_{r_{s-}}^{r_{s+}}
\end{equation}
parameterizes the amplitude and phase of the current sheet flowing at the resonant surface. Matching the solutions
in the so-called {\em inner region}\/ (i.e., the region of the plasma in the immediate vicinity of the resonant surface)
and the so-called {\em outer region}\/ (i.e., everywhere in the plasma other than the inner region) with the help
of Equations~(\ref{e99}), (\ref{e105}), and (\ref{e106}), we obtain
\begin{equation}\label{e110}
{\mit\Delta\Psi}_s = E_{ss}\,{\mit\Psi}_s,
\end{equation}
where $E_{ss}= {\mit\Delta}_{s+}-{\mit\Delta}_{s-}$ is a real dimensionless quantity that is known as the {\em tearing stability index}\/ (Furth, et al.\ 1963).

@@

\subsection{Solution of Plasma Angular Equations of Motion}
In general, the perturbed angular velocity profiles, ${\mit\Delta\Omega}_\theta(r,t)$ and ${\mit\Delta\Omega}_\varphi(r,t)$, are
localized in the vicinity of the resonant surface (Fitzpatrick 1993). Hence, it is reasonable to express  the perturbed angular
equations of motion, (\ref{e182}) and (\ref{e185}), in the simplified  forms
\begin{align}\label{e188}
4\pi^2\,R_0\left[(1+2\,q_s^{\,2})\,\rho_s\,r^{\,3}\,\frac{\partial {\mit\Delta\Omega}_\theta}{\partial t}+ \frac{\rho_s}{\tau_{\theta\,s}}\,r^{\,3}\,{\mit\Delta\Omega}_\theta-\mu_{\perp\,s}\frac{\partial}{\partial r}\!\left(r^{\,3}\,\frac{\partial{\mit\Delta\Omega}_\theta}{\partial r}\right)\right]
&=T_{\theta\,s}\,\delta(r-r_s),\\[0.5ex]
4\pi^2\,R_0^{\,3}\left[\rho_s\,r\,\frac{\partial {\mit\Delta\Omega}_\varphi}{\partial t}-\mu_{\perp\,s}\frac{\partial}{\partial r}\!\left(r\,\frac{\partial{\mit\Delta\Omega}_\varphi}{\partial r}\right)\right]
=T_{\varphi\,s}\,\delta(r-r_s),\label{e189}
\end{align}
where $q_s=q(r_s)$, $\rho_s=\rho(r_s)$, $\tau_{\theta\,s}=\tau_\theta(r_s)$, and $\mu_{\perp\,s}=\mu_\perp(r_s)$.

@@

Equations~(\ref{e220})--(\ref{e223}) are known collectively as the {\em drift-MHD equations},\footnote{The 'drift' appellation  merely indicates that we are taking diamagnetic flows into account.} and contain both shear-Alfv\'{e}n
and compressible-Alfv\'{e}n wave dynamics. The system of equations is completed by a subset of Maxwell's equations:
\begin{align}
\nabla\cdot{\bf B} &= 0,\\[0.5ex]
\nabla\times {\bf E} &= - \frac{\partial {\bf B}}{\partial t},\\[0.5ex]
\mu_0\,{\bf j} &= \nabla \times {\bf B}.\label{e226}
\end{align}

@@

Strictly speaking, the analysis
presented in this section is only valid when the ion gyroradius is smaller than the typical variation lengthscale in the resonant layer. 
If this is not the case then it is necessary to adopt a so-called {\em gyrofluid}\/ approach in which the ion response is
averaged over the ion gyro-orbits (Cowley, Kulsrud \& Hahm 1986; Hammett, Dorland \& Perkins 1992; Waelbroeck, Hazeltine \& Morrison 2009). Of course, if the ions are cold (i.e., $T_i\ll T_e$) then the ion gyroradius tends to zero,  which means that the
analysis is valid. 
Note, however, that the ion sound radius is finite even when the ions are cold. In the following, we shall assume that the
analysis is valid even when the ions are hot.

@@

In conventional tokamak plasmas, the Lundquist number, $S$, which is the nominal ratio of the plasma inertia to the   resistive diffusion  terms in the plasma Ohm's law, is very much greater than unity. In fact, $S$ typically exceeds $10^8$ (Wesson 2011). However, a resonant layer is
characterized by a balance between plasma inertia and resistive diffusion (Furth, et al.\ 1963). Such a balance is
only possible if the layer is very narrow in the radial direction (because a narrow layer enhances radial derivatives, and, thereby, enhances
resistive diffusion). Let us define the stretched radial variable (Ara, et al.\ 1978)
\begin{equation}
X = S^{1/3}\,\hat{x}.
\end{equation}
Assuming that $X\sim{\cal O}(1)$ in the layer (i.e., assuming that the layer thickness is of order $S^{-1/3}\,r_s$),
and making use of the fact that $S\gg 1$, we deduce that $\hat{\nabla}^2\simeq \hat{\nabla}_\perp^2\simeq S^{\,2/3}\,d^{\,2}/dX^2$.
Hence, the linear layer equations, (\ref{e301})--(\ref{e304}), reduce to (Cole \& Fitzpatrick 2006)
\begin{align}\label{e306}
-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_{e}]\,\tilde{\psi}&= - {\rm i}\,X\left[\tilde{\phi}-(1+\lambda_e)\,\tilde{N}\right] + \frac{d^{\,2}\tilde{\psi}}{d X^2},\\[0.5ex]
-{\rm i}\,(Q-Q_E -Q_{i})\,\frac{d^{\,2}\tilde{\phi}}{dX^2}&= - {\rm i}\,X\,\frac{d^{\,2}\tilde{\psi}}{dX^2}+ P_\varphi\,\frac{d^{\,4}}{dX^4}\!\left(\tilde{\phi} + \frac{\tilde{N}}{\tau}\right),\\[0.5ex]
-{\rm i}\,(Q-Q_E)\tilde{N} &= - {\rm i}\,Q_{e}\,\tilde{\phi} - {\rm i} \,D^{2}\,X\,\frac{d^{\,2}\tilde{\psi}}{dX^{2}}
 - {\rm i}\,c_\beta^{\,2}\,X\,\tilde{V} + P_\perp\,\frac{d^{\,2} \tilde{N}}{dX^{2}},\label{e308}\\[0.5ex]
 -{\rm i}\,(Q-Q_E)\,\tilde{V} &= {\rm i}\,Q_{e}\,\tilde{\psi} - {\rm i}\,X\,\tilde{N} + P_\varphi\,\frac{d^{\,2}\tilde{V}}{dX^{2}}.\label{e309}
\end{align}
Here, $Q=S^{\,1/3}\,\omega\,\tau_H$, $Q_E = S^{\,1/3}\,\omega_E\,\tau_H$, $Q_{e,i} = S^{\,1/3}\,\omega_{\ast\,e,i}\,\tau_H$,
and $D = S^{\,1/3}\,[\tau/(1+\tau)]^{1/2}\,\hat{d}_s$. Note that the terms involving parallel transport have dropped out of
Equation~(\ref{e308}). The neglect of these terms is justified provided that the linear layer width is much less than the critical value,
\begin{equation}
\delta_c = \left(\frac{1}{\epsilon\,s\,n}\sqrt{\frac{D_\perp}{D_\parallel}}\right)^{1/2}_{r_s} r_s,
\end{equation}
below which parallel transport is unable to constrain the electron number density (and, by implication, the electron and
ion temperatures) to be constant along magnetic field-lines (Fitzpatrick 1995). This condition is well satisfied in conventional
tokamak plasmas. 
Note that all of the terms appearing in the 
normalized layer equations, (\ref{e306})--(\ref{e309}), are assumed to be roughly the same order of magnitude.

@@

\subsection{Constant-$\psi$ Approximation}\label{s5.6}
Let us suppose that there are two layers in $p$ space. In the small-$p$ layer, suppose that Equation~(\ref{e320}) reduces to
\begin{equation}
\frac{d}{dp}\!\left[\frac{p^2}{-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e] + p^2}\,\frac{dY}{dp}\right]\simeq 0
\end{equation}
when $p\sim Q^{1/2}$. Integrating directly, we find that
\begin{equation}
Y(p) \simeq Y_0 \left[\frac{\hat{\mit\Delta}}{\pi}\left(\frac{1}{p} + \frac{p}{{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e]}\right) + 1 + {\cal O}(p^2)\right]
\end{equation}
for $p\lesssim {\cal O}(Q^{1/2})$, where use has been made of Equation~(\ref{e325}). The two-layer  approximation is
equivalent to the well-known {\em constant-$\psi$ approximation}\/ (Furth, et al.\ 1963).

@@

\subsection{Constant-$\psi$ Linear Response Regimes}\label{scp}
Suppose that $Q\gg P\,p^2$ and $D^2\,p^2 \ll 1$. It follows that $\nu=1/4$ and
\begin{equation}
C = \frac{[-{\rm i}\,(Q-Q_E)]\,[-{\rm i}\,(Q-Q_E-Q_i)]}{-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e]}.
\end{equation}
Hence, we deduce that
\begin{equation}\label{ri}
\hat{\mit\Delta} = \frac{2\pi\,{\mit\Gamma}(3/4)}{{\mit\Gamma}(1/4)}\,(-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e])^{3/4}\,[-{\rm i}\,(Q-Q_E)]^{1/4}\,[-{\rm i}\,(Q-Q_E-Q_i)]^{1/4}.
\end{equation}
This response regime is known as the {\em resistive-inertial regime}, because the layer response is dominated by
resistivity and ion inertia (Furth, et al.\ 1963; Ara, et al.\ 1976; Fitzpatrick 1998). The characteristic layer width is $p_\ast \sim Q^{-1/4}$,
which implies that the regime is valid when $P\ll Q^{3/2}$, $Q\gg D^4$, and $Q\ll 1$.

@@

Suppose that the resonant layer lies in a constant-$\psi$ response regime. According to Equation~(\ref{e295}),
the perturbed magnetic flux in the layer takes the form:
\begin{equation}
\hat{\psi}(\hat{x},\zeta,t) = \frac{\hat{x}^{\,2} }{2\,\hat{L}_s}+ \tilde{\psi}\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},
\end{equation}
where $\tilde{\psi}$ is a complex constant. Of course, the physical flux is the real part of the previous expression; that is,
\begin{equation}\label{e352a}
\hat{\psi}(\hat{x},\xi)= \frac{\hat{x}^{\,2} }{2\,\hat{L}_s}+ |\tilde{\psi}|\,\cos\xi,
\end{equation}
where
\begin{equation}
\xi = \zeta +{\rm arg}(\tilde{\psi})-\hat{\omega}\,\hat{t} = m\,\theta-n\,\varphi +{\rm arg}(\tilde{\psi}) -\omega\,t.
\end{equation}
Now, it is clear from Equation~(\ref{e279}) that ${\bf B}\cdot\hat{\nabla}\hat{\psi}=0$. In other words, the contours of
$\hat{\psi}(\hat{x},\xi)$ map out the perturbed magnetic flux-surfaces in the vicinity of the resonant layer. 
Let $\hat{W} = 4\,(|\tilde{\psi}|\,\hat{L}_s)^{1/2}$. Equation~(\ref{e352a}) can be written
\begin{equation}
\frac{\psi(\hat{x},\xi)}{|\tilde{\psi}|} = 8\left(\frac{\hat{x}}{\hat{W}}\right)^2 + \cos\xi.
\end{equation}
Figure~\ref{f2a} shows the contours of $\psi(\hat{x},\xi)$ specified by the previous equation. It can be seen that the
tearing mode has reconnected magnetic field-lines to produce a {\em helical magnetic island chain}\/ centered on the resonant
surface. The chain has $m$ periods in the poloidal direction, $n$ periods in the toroidal direction, and propagates at the
helical phase velocity $\omega$. The maximum full (as opposed to half) radial width of the magnetic separatrix is 
\begin{equation}\label{e355s}
W =  \hat{W}\,r_s= 4\,(|\tilde{\psi}|\,\hat{L}_s)^{1/2}\,r_s = 4\left(\frac{q_s}{s_s}\,|\hat{\mit\Psi}_s|\right)^{1/2}R_0,
\end{equation}
where ${\mit\Psi}_s = r_s\,B_z\,\tilde{\psi}$ is the reconnected magnetic flux defined in Equation~(\ref{e99}), and
$\hat{\mit\Psi}_s= {\mit\Psi}_s/(R_0\,B_z)$.

@@

The absence of nonconstant-$\psi$ response regimes in Figure~\ref{f3}
should come as no surprise. As we saw in Section~\ref{dis1}, nonconstant-$\psi$ resonant layers
are characterized by $\hat{\mit\Delta}\sim\hat{\delta}_\ast^{-1}= S^{-1/3}\,(r_s/\delta)$, where $\delta$
is the layer thickness. Hence, according to Equation~(\ref{e352}), asymptotic matching of such a
layer to the outer solution is only possible if $|E_{ss}|\sim r_s/\delta \gg 1$ (given that resonant
layers in tokamak plasmas are invariably very thin compared to the minor radius of the plasma). However,
low-$m$ tearing modes in conventional tokamak plasmas are characterized by $|E_{ss}|\sim 1$ rather than
$|E_{ss}|\gg 1$. (As before, we are neglecting $m=1$ modes, which are characterized by
$|E_{ss}|\gg 1$, because they are not really tearing modes.)

>>>>

If we assume that $\delta_e\sim \zeta_e\sim \mu\ll 1$ then the dominant term in the electron energy conservation equation
(\ref{e3.151c}) yields
\begin{equation}
\nabla\cdot{\bf q}_{T_e}\simeq \nabla\cdot(-\kappa_{\parallel\,e}\,\nabla_\parallel T_e)\simeq 0,
\end{equation}
which implies that 
\begin{equation}
{\bf B}\cdot\nabla T_e=0. \label{e75}
\end{equation}
In other words, the parallel electron heat conductivity in a tokamak plasma is usually
sufficiently large that it forces the electron temperature to be constant on magnetic flux-surfaces. 
Likewise, if we assume that $\delta_i\sim \zeta_i\sim \mu\ll 1$ then the dominant term in the ion energy conservation equation
(\ref{e3.152c}) yields
\begin{equation}
\nabla\cdot{\bf q}_{T_i}\simeq \nabla\cdot(-\kappa_{\parallel\,i}\,\nabla_\parallel T_i)\simeq 0,
\end{equation}
which implies that 
\begin{equation}
{\bf B}\cdot\nabla T_i=0.\label{e77}
\end{equation}
In other words, the parallel ion heat conductivity in a tokamak plasma is usually
sufficiently large that it forces the ion temperature to be constant on magnetic flux-surfaces.

 The dominant terms in the electron and ion momentum conservation equations,
(\ref{e3.151b}) and (\ref{e3.152b}), respectively, yield
\begin{align}\label{e78}
\nabla p_e + n\,e\,({\bf E}+{\bf V}_e\times {\bf B}) &\simeq 0,\\[0.5ex]
\nabla p_i - n\,e\,({\bf E}+{\bf V}_i\times {\bf B}) &\simeq 0.\label{e79}
\end{align}
The sum of the preceding two equations gives
\begin{equation}\label{e80}
 {\bf j}\times {\bf B} \simeq\nabla p,
\end{equation}
where $p=p_e+p_i$. In other words, a conventional tokamak plasma exists in a state in which the magnetic force density exactly balances the  total scalar pressure force density. 
It follows from the previous equation that 
\begin{equation}\label{e81}
{\bf B}\cdot\nabla p\simeq 0.
\end{equation}
In other words, the total pressure in a conventional tokamak plasma is constant on magnetic flux-surfaces. 
 However, given that $p=n\,(T_e+T_i)$, and making use of Eqs.~(\ref{e75}) and (\ref{e77}), we deduce that 
\begin{equation}
{\bf B}\cdot\nabla n\simeq 0.
\end{equation}
In other words, the electron number density in a conventional tokamak plasma is also  constant on
magnetic flux-surfaces.

Let us make the reasonable assumption that Equation~(\ref{e81}) implies that ${\bf B}\cdot\nabla p_e={\bf B}\cdot\nabla
p_i=0$. In other words, the electron and ion pressures are separately constant on magnetic flux-surfaces. If we assume that, to lowest order, ${\bf E}=-\nabla{\mit\Phi}$, where ${\mit\Phi}$ is the electrostatic potential, then Equations~(\ref{e78})
and (\ref{e79}) yield
\begin{equation}
{\bf B}\cdot\nabla{\mit\Phi}=0.
\end{equation}
In other words,  the electrostatic potential is also constant on magnetic flux-surfaces.

@@

\subsection{Solution in Presence of Perfectly-Conducting Wall}\label{perfect}
Suppose that the plasma occupies the region $0\leq r\leq a$, where $a$ is the plasma {\em minor radius}. (By definition, $a\ll R_0$ in a large aspect-ratio tokamak.( It follows that
$J_z(r)=0$ for $r>a$. Let the plasma be surrounded by a concentric, rigid, radially-thin, perfectly-conducting, wall of radius $r_w>a$. 
(In most circumstances, the wall represents the metal vacuum vessel that surrounds  the plasma.) An appropriate
physical solution of the cylindrical tearing mode equation takes the separable form
\begin{equation}\label{e99}
\psi(r,t)= {\mit\Psi}_s(t)\,\hat{\psi}_s(r),
\end{equation}
where the real function $\hat{\psi}_s(r)$ is a solution of 
\begin{equation}\label{e100}
\frac{d^2\hat{\psi}_s}{dr^2} + \frac{1}{r}\,\frac{d\hat{\psi}_s}{dr}-\frac{m^2}{r^2}\,\hat{\psi}_s - \frac{J_z'\,\hat{\psi}_s}{r\,(1/q-n/m)}= 0
\end{equation}
that satisfies
\begin{align}\label{e101}
\hat{\psi}_s(0) &= 0,\\[0.5ex]
\hat{\psi}_s(r_s) &= 1,\label{e102}\\[0.5ex]
\hat{\psi}_s(r\geq r_w) &= 0.\label{e103}
\end{align}

@@

The complex quantity
\begin{equation}\label{e109}
{\mit\Delta\Psi}_s = \left[r\,\frac{\partial \psi}{\partial r}\right]_{r_{s-}}^{r_{s+}}
\end{equation}
parameterizes the amplitude and phase of the current sheet flowing at the resonant surface. Matching the solutions
in the so-called {\em inner region}\/ (i.e., the region of the plasma in the immediate vicinity of the resonant surface)
and the so-called {\em outer region}\/ (i.e., everywhere in the plasma other than the inner region) with the help
of Equations~(\ref{e99}), (\ref{e105}), and (\ref{e106}), we obtain
\begin{equation}\label{e110}
{\mit\Delta\Psi}_s = E_{ss}\,{\mit\Psi}_s,
\end{equation}
where $E_{ss}= {\mit\Delta}_{s+}-{\mit\Delta}_{s-}$ is a real dimensionless quantity that is known as the {\em tearing stability index}\/ (Furth, Killeen \& Rosenbluth 1963).

@@

\subsection{Solution of Plasma Angular Equations of Motion}
In most situations of interest, the perturbed angular velocity profiles, ${\mit\Delta\Omega}_\theta(r,t)$ and ${\mit\Delta\Omega}_\varphi(r,t)$, are
localized in the vicinity of the resonant surface (Fitzpatrick 1993). Hence, it is reasonable to express  the perturbed angular
equations of motion, (\ref{e182}) and (\ref{e185}), in the simplified  forms
\begin{align}\label{e188}
4\pi^2\,R_0\left[(1+2\,q_s^{\,2})\,\rho_s\,r^{\,3}\,\frac{\partial {\mit\Delta\Omega}_\theta}{\partial t}+ \frac{\rho_s}{\tau_{\theta\,s}}\,r^{\,3}\,{\mit\Delta\Omega}_\theta-\mu_{\perp\,s}\frac{\partial}{\partial r}\!\left(r^{\,3}\,\frac{\partial{\mit\Delta\Omega}_\theta}{\partial r}\right)\right]
&=T_{\theta\,s}\,\delta(r-r_s),\\[0.5ex]
4\pi^2\,R_0^{\,3}\left[\rho_s\,r\,\frac{\partial {\mit\Delta\Omega}_\varphi}{\partial t}-\mu_{\perp\,s}\frac{\partial}{\partial r}\!\left(r\,\frac{\partial{\mit\Delta\Omega}_\varphi}{\partial r}\right)\right]
=T_{\varphi\,s}\,\delta(r-r_s),\label{e189}
\end{align}
where $q_s=q(r_s)$, $\rho_s=\rho(r_s)$, $\tau_{\theta\,s}=\tau_\theta(r_s)$, and $\mu_{\perp\,s}=\mu_\perp(r_s)$.

@@

Equations~(\ref{e220})--(\ref{e223}) are known collectively as the {\em drift-MHD equations},\footnote{The  `drift' appellation  merely indicates that we are taking diamagnetic flows into account.} and contain both shear-Alfv\'{e}n
and compressible-Alfv\'{e}n wave dynamics. The system of equations is completed by a subset of Maxwell's equations:
\begin{align}
\nabla\cdot{\bf B} &= 0,\\[0.5ex]
\nabla\times {\bf E} &= - \frac{\partial {\bf B}}{\partial t},\\[0.5ex]
\mu_0\,{\bf j} &= \nabla \times {\bf B}.\label{e226}
\end{align}

@@

Strictly speaking, the four-field fluid model 
presented in this section is only valid when the ion gyroradius is smaller than the typical variation lengthscale in the resonant layer. 
If this is not the case then it is necessary to adopt a so-called {\em gyrofluid}\/ approach in which the ion response is
averaged over the ion gyro-orbits (Cowley, Kulsrud \& Hahm 1986; Hammett, Dorland \& Perkins 1992; Waelbroeck, Hazeltine \& Morrison 2009). Of course, if the ions are cold (i.e., $T_i\ll T_e$) then the ion gyroradius tends to zero,  which means that the
analysis is valid. 
Note, however, that the ion sound radius is finite even when the ions are cold. In the following, we shall assume that our model remains valid even when the ions are hot.

@@

In conventional tokamak plasmas, the Lundquist number, $S$, which is the nominal ratio of the plasma inertia to the   resistive diffusion  terms in the plasma Ohm's law, is very much greater than unity. In fact, $S$ typically exceeds $10^8$ (Wesson 2011). However, a resonant layer is
characterized by a balance between plasma inertia and resistive diffusion (Furth, Killeen \& Rosenbluth 1963). Such a balance is
only possible if the layer is very narrow in the radial direction (because a narrow layer enhances radial derivatives, and, thereby, enhances
resistive diffusion). Let us define the stretched radial variable (Ara, et al.\ 1978)
\begin{equation}
X = S^{1/3}\,\hat{x}.
\end{equation}
Assuming that $X\sim{\cal O}(1)$ in the layer (i.e., assuming that the layer thickness is of order $S^{-1/3}\,r_s$),
and making use of the fact that $S\gg 1$, we deduce that $\hat{\nabla}^2\simeq \hat{\nabla}_\perp^2\simeq S^{\,2/3}\,d^{\,2}/dX^2$.
Hence, the linear layer equations, (\ref{e301})--(\ref{e304}), reduce to (Cole \& Fitzpatrick 2006)
\begin{align}\label{e306}
-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_{e}]\,\tilde{\psi}&= - {\rm i}\,X\left[\tilde{\phi}-(1+\lambda_e)\,\tilde{N}\right] + \frac{d^{\,2}\tilde{\psi}}{d X^2},\\[0.5ex]
-{\rm i}\,(Q-Q_E -Q_{i})\,\frac{d^{\,2}\tilde{\phi}}{dX^2}&= - {\rm i}\,X\,\frac{d^{\,2}\tilde{\psi}}{dX^2}+ P_\varphi\,\frac{d^{\,4}}{dX^4}\!\left(\tilde{\phi} + \frac{\tilde{N}}{\tau}\right),\\[0.5ex]
-{\rm i}\,(Q-Q_E)\tilde{N} &= - {\rm i}\,Q_{e}\,\tilde{\phi} - {\rm i} \,D^{2}\,X\,\frac{d^{\,2}\tilde{\psi}}{dX^{2}}
 - {\rm i}\,c_\beta^{\,2}\,X\,\tilde{V} + P_\perp\,\frac{d^{\,2} \tilde{N}}{dX^{2}},\label{e308}\\[0.5ex]
 -{\rm i}\,(Q-Q_E)\,\tilde{V} &= {\rm i}\,Q_{e}\,\tilde{\psi} - {\rm i}\,X\,\tilde{N} + P_\varphi\,\frac{d^{\,2}\tilde{V}}{dX^{2}}.\label{e309}
\end{align}
Here, $Q=S^{\,1/3}\,\omega\,\tau_H$, $Q_E = S^{\,1/3}\,\omega_E\,\tau_H$, $Q_{e,i} = S^{\,1/3}\,\omega_{\ast\,e,i}\,\tau_H$,
and $D = S^{\,1/3}\,[\tau/(1+\tau)]^{1/2}\,\hat{d}_s$. Note that the terms involving parallel transport have dropped out of
Equation~(\ref{e308}). The neglect of these terms is justified provided that the linear layer width is much less than the critical value,
\begin{equation}
\delta_c = \left(\frac{1}{\epsilon\,s\,n}\sqrt{\frac{D_\perp}{D_\parallel}}\right)^{1/2}_{r_s} r_s,
\end{equation}
below which parallel transport is unable to constrain the electron number density (and, by implication, the electron and
ion temperatures) to be constant along magnetic field-lines (Fitzpatrick 1995). This condition is well satisfied in conventional
tokamak plasmas. 
Note that all of the terms appearing in the 
normalized layer equations, (\ref{e306})--(\ref{e309}), are assumed to be roughly the same order of magnitude.

@@

\subsection{Constant-$\psi$ Approximation}\label{s5.6}
Let us suppose that there are two layers in $p$ space. In the small-$p$ layer, suppose that Equation~(\ref{e320}) reduces to
\begin{equation}
\frac{d}{dp}\!\left[\frac{p^2}{-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e] + p^2}\,\frac{dY}{dp}\right]\simeq 0
\end{equation}
when $p\sim Q^{1/2}$. Integrating directly, we find that
\begin{equation}
Y(p) \simeq Y_0 \left[\frac{\hat{\mit\Delta}}{\pi}\left(\frac{1}{p} + \frac{p}{{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e]}\right) + 1 + {\cal O}(p^2)\right]
\end{equation}
for $p\lesssim {\cal O}(Q^{1/2})$, where use has been made of Equation~(\ref{e325}). The two-layer  approximation is
equivalent to the well-known {\em constant-$\psi$ approximation}\/ (Furth, Killeen \& Rosenbluth 1963).

@@

\subsection{Constant-$\psi$ Linear Response Regimes}\label{scp}
Suppose that $Q\gg P\,p^2$ and $D^2\,p^2 \ll 1$. It follows that $\nu=1/4$ and
\begin{equation}
C = \frac{[-{\rm i}\,(Q-Q_E)]\,[-{\rm i}\,(Q-Q_E-Q_i)]}{-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e]}.
\end{equation}
Hence, we deduce that
\begin{equation}\label{ri}
\hat{\mit\Delta} = \frac{2\pi\,{\mit\Gamma}(3/4)}{{\mit\Gamma}(1/4)}\,(-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e])^{3/4}\,[-{\rm i}\,(Q-Q_E)]^{1/4}\,[-{\rm i}\,(Q-Q_E-Q_i)]^{1/4}.
\end{equation}
This response regime is known as the {\em resistive-inertial regime}, because the layer response is dominated by
resistivity and ion inertia (Furth, Killeen \& Rosenbluth 1963; Ara, et al.\ 1976; Fitzpatrick 1998). The characteristic layer width is $p_\ast \sim Q^{-1/4}$,
which implies that the regime is valid when $P\ll Q^{3/2}$, $Q\gg D^4$, and $Q\ll 1$.

@@

Suppose that the resonant layer lies in a constant-$\psi$ response regime. According to Equation~(\ref{e295}),
the perturbed magnetic flux in the layer takes the form:
\begin{equation}
\hat{\psi}(\hat{x},\zeta,t) = \frac{\hat{x}^{\,2} }{2\,\hat{L}_s}+ \tilde{\psi}\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},
\end{equation}
where $\tilde{\psi}$ is a complex constant. Of course, the physical flux is the real part of the previous expression; that is,
\begin{equation}\label{e352a}
\hat{\psi}(\hat{x},\xi)= \frac{\hat{x}^{\,2} }{2\,\hat{L}_s}+ |\tilde{\psi}|\,\cos\xi,
\end{equation}
where
\begin{equation}
\xi = \zeta +{\rm arg}(\tilde{\psi})-\hat{\omega}\,\hat{t} = m\,\theta-n\,\varphi +{\rm arg}(\tilde{\psi}) -\omega\,t.
\end{equation}
Now, it is clear from Equation~(\ref{e279}) that ${\bf B}\cdot\hat{\nabla}\hat{\psi}=0$. In other words, the contours of
$\hat{\psi}(\hat{x},\xi)$ map out the perturbed magnetic flux-surfaces in the vicinity of the resonant layer. 
Let $\hat{W} = 4\,(|\tilde{\psi}|\,\hat{L}_s)^{1/2}$. Equation~(\ref{e352a}) can be written
\begin{equation}
\frac{\psi(\hat{x},\xi)}{|\tilde{\psi}|} = 8\left(\frac{\hat{x}}{\hat{W}}\right)^2 + \cos\xi.
\end{equation}
Figure~\ref{f2a} shows the contours of $\psi(\hat{x},\xi)$ specified by the previous equation. It can be seen that the
tearing mode has reconnected magnetic field-lines to produce a {\em helical magnetic island chain}\/ centered on the resonant
surface ($\hat{x}=0$). The chain has $m$ periods in the poloidal direction, $n$ periods in the toroidal direction, and propagates at the
helical phase velocity $\omega$. The maximum full (as opposed to half) radial width of the magnetic separatrix is 
\begin{equation}\label{e355s}
W =  \hat{W}\,r_s= 4\,(|\tilde{\psi}|\,\hat{L}_s)^{1/2}\,r_s = 4\left(\frac{q_s}{s_s}\,|\hat{\mit\Psi}_s|\right)^{1/2}R_0,
\end{equation}
where ${\mit\Psi}_s = r_s\,B_z\,\tilde{\psi}$ is the reconnected magnetic flux defined in Equation~(\ref{e99}), and
$\hat{\mit\Psi}_s= {\mit\Psi}_s/(R_0\,B_z)$.

@@

The absence of nonconstant-$\psi$ response regimes in Figure~\ref{f3}
should come as no surprise. As we saw in Section~\ref{dis1}, nonconstant-$\psi$ resonant layers
are characterized by $\hat{\mit\Delta}\sim\hat{\delta}_\ast^{-1}= S^{-1/3}\,(r_s/\delta)$, where $\delta$
is the layer thickness. Hence, according to Equation~(\ref{e352}), asymptotic matching of such a
layer to the outer solution is only possible if $|E_{ss}|\sim r_s/\delta \gg 1$ (given that resonant
layers in tokamak plasmas are invariably very thin compared to the minor radius of the plasma). However,
low-$m$ tearing modes in conventional tokamak plasmas are characterized by $|E_{ss}|\sim {\cal O}(1)$ rather than
$|E_{ss}|\gg 1$. (As before, we are neglecting $m=1$ modes, which are characterized by
$|E_{ss}|\gg 1$, because they are not really tearing modes.)
